---
title: "STAT 527 HW 1"
author: "Satoshi Ido (ID: 34788706)"
date: 23 January 2023
output: pdf_document
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    warning = FALSE,
    message = FALSE)
```

1\
(a)Evaluate the following expressions.
```{r message = FALSE}
a <- (93^2 - 164) / (46^3 + 189)
b <- 376 - (23^2) / 4
c <- (59 + 48^2) / ((-9) + 22^2) - (-16 + 55^2) / (13 + 29^2)
d <- 18^4 - 16^3 + 14^2 - 12
c(a, b, c, d)
```
1\
(b)Evaluate 3^x for x = 1, 2, . . . , 20 and store the values in a vector. Print the vector with the function print(). Report the length of the vector with the function length().
```{r message = FALSE}
x <- 1:20
print(3^x)
length(3^x)
```

2\
(10 points) If you throw a baseball at an angle of 45◦, at an initial velocity of 75 mph, while standing on a level field, the ball's horizontal distance x traveled after t seconds is described (neglecting air resistance) by the following equation from Newtonian physics: x = 27.12t
Furthermore, the height above the ground after t seconds, assuming the ball was initially released at a height of 5 ft, is described by y = 1.524 + 19.71t - 4.905t^2.
The equations have been calibrated to give the distance x and height y in meters. The ball will hit the ground after about 4.09 seconds. 
Calculate a vector (say, x) of baseball distances for a range of values of t from 0 to 4.09. Calculate a vector of baseball heights (say, y) for the same collection of times. Make a plot of x (horizontal axis) and y (vertical axis). 
Read from the graph of the ball's trajectory how high and how far, approximately, the ball will travel.
```{r message = FALSE, out.width = "50%"}
t <- 0:4.09
x <- 27.12 * t
y <- 1.524 + 19.71 * t - 4.905 * t^2
plot(x, y)

# From the plot, the ball goes up to a maximum of 21 ft and flies far up to 130 ft.
```

3\
(10 points) To decrease the use of insecticides in agriculture, predator insects are often released to combat insect pests. Coccinellids (lady beetles), in particular, have a voracious appetite for aphids. 
In a recent study (Pervez and Omkar 2005), entomologists looked at the suitability of using coccinellids to control a particular aphid, Myzus persicae (common name is the ”green peach aphid”), a serious pest of many fruit and vegetable crops. 
In the study, the entomologists experimentally ascertained aphid kill rates for three different species of coccinellids:
Enter the data columns above into vectors, giving them descriptive names. For each type of coccinellid, use R to construct a scatterplot (type = “p”) of the feeding rate of the coccinellid versus aphid density.
```{r message = FALSE, out.width = "50%"}
# vectors data
cs <- c(21, 37, 65, 102, 125, 141, 154, 164, 170, 177)
ct <- c(21, 37, 60, 90, 109, 120, 129, 136, 140, 143)
pd <- c(15, 26, 42, 59, 69, 74, 79, 83, 85, 82)
# plot for each type of coccinellid
## vector data of aphid density
ad <- c(25, 50, 100, 200, 300, 400, 500, 600, 700, 800)
## plot
plot(cs, ad, type = "p", main = "the Cheilomenes sexmaculata vs Aphid density")

plot(ct, ad, type = "p", main = "the Coccinella transversalis vs Aphid density")

plot(pd, ad, type = "p", main = "the Propylea dissecta vs Aphid density")
```