---
title: "STAT 656 HW 2"
author: "Satoshi Ido (ID: 34788706)"
date: 21 September 2023
output: pdf_document
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    warning = FALSE,
    message = FALSE)
```

```{r}
library("ggplot2")
library("dplyr")
library("tidyr")
library("MASS")
library("zoom")
library("lmtest")
library("fs")
library("moments")
library("rstan")
library("bayesplot")
library("StanHeaders")
```

set options to speed up the calculations
```{r}
# to avoid recompilation of unchanged Stan programs, we recommend calling
rstan_options(auto_write = TRUE)
# for execution on a local, multicore CPU with excess RAM we recommend calling
options(mc.cores = parallel::detectCores())
# set the size of the plots
options(repr.plot.width = 12, repr.plot.height = 6)

# options(bayesplot::theme_default())
bayesplot_theme_set(theme_default(base_size = 24, base_family = "sans"))
```

```{r}
# Create the input_dir (input directory)
current_note_path <- getwd()
INPUT_DIR <- file.path(current_note_path, "656/hw/hw2/data")

# If INPUT_DIR has not been created yet, create it
if (!dir.exists(INPUT_DIR)) {
  dir.create(INPUT_DIR)
}

# Create the output_dir (output directory)
OUTPUT_DIR <- file.path(current_note_path, "656/hw/hw2/outputs")

# If OUTPUT_DIR has not been created yet, create it
if (!dir.exists(OUTPUT_DIR)) {
  dir.create(OUTPUT_DIR)
}

# Read CSV files using a function to specify the directory automatically
read_csv <- function(name, ...) {
  path <- file.path(INPUT_DIR, paste0(name, ".csv"))
  print(paste("Load:", path))
  return(read.csv(path, ...))
}
```

# first design selection
My first approach is to run the eperiments with 24 wells initially, followed by 48 wells in the second experiments. 
The reason for this is that I want to see if the 24 wells are enough to get the information about the mean and variance of the population so that I can set a reasonable prior for the second experiment.
I will assign pairs of concentrations for the two chemical modulators to each well in the manner as below. The main focus is to see how the effect of the one modulator changes depending on the concentration of the other modulator.
Since I have little idea of the effect of the modulators, I will assign the concentrations of the modulators somewhat randomly.
```{r}
design1 <- matrix(nrow=24, ncol=2)
# assign the concentration of the modulators randomly
## set a seed to reproduce the same result
set.seed(49)
modA <- rep(seq(0, 75, by = 15), each = 4)
modB <- rep(seq(0, 30, by = 10), times = 6)
design1 <- cbind(modA, modB)
# write a table and save it to csv
write.table(design1, file = file.path(OUTPUT_DIR, "design1.csv"), sep = ",", col.names = F, row.names = F)
```
