---
title: "STAT 526 HW 1"
author: "Satoshi Ido (ID: 34788706)"
date: 23 January 2023
output: pdf_document
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    warning = FALSE,
    message = FALSE)
```

# Summary


# Analysis


# Appendix

Library import
```{r}
library("tidyverse")
library("dplyr")
library("car")
library("ggplot2")
library("reshape2")
library("caTools")
library("Rcmdr")
library("MASS")
```

data import
```{r message = FALSE}
data(Cars93, package = "MASS")
```

data display
```{r message = FALSE}
View(Cars93)
# number of rows
nrow(Cars93)
# data type of each column
str(Cars93)
# if there is null values in the dataset
# is.na(Cars93)
# name of the columns in the dataset
names(Cars93)

```

create the response variable "MPG.avg" by averaging "MPG.city" and "MPG.highway"
```{r message = FALSE}
df <- Cars93 %>%
        mutate(MPG.avg = (Cars93$MPG.highway + Cars93$MPG.city) / 2)
```

check the distribution of response (histogram, qq-plot) → box-cox
```{r meesage = FALSE, out.width = "50%"}
plot(density(df$MPG.avg))
qqnorm(df$MPG.avg, pch = 1, frame = FALSE)
qqline(df$MPG.avg, col = "steelblue", lwd = 2)
```


box-cox
```{r message = FALSE, out.width = "50%"}
## there is no 0 in MPG.avg
min(df$MPG.avg); max(df$MPG.avg)

df2 <- df %>%
        na.omit() %>%
        mutate(Cylinders = as.numeric(.$Cylinders)) %>%
        subset(select = -c(MPG.city, MPG.highway, Min.Price, Max.Price, Manufacturer, Model, Make)) #%>%
        # remove the non-numerical variables
        # https://bit.ly/3ZXRAMH
        #.[, colnames(.)[!grepl("factor|logical|character", sapply(., class))]]

# boxcox
m1 <- lm(MPG.avg ~ ., data = df2)
bc <- boxcox(m1)
# lambda = -2
lambda <- bc$x[which.max(bc$y)]
ggplot(data = df2, aes(x = (MPG.avg^lambda -1)/lambda)) + geom_histogram(fill = "blue", bins = 15)

# update MPG.avg
df2$MPG.avg <- ((df2$MPG.avg^lambda - 1) / lambda)
m1_new <- lm(MPG.avg ~ ., data = df2)

# check
# summary(m1_new)
# par(mfrow = c(2, 2))
# plot(m1_new)

# plot again
plot(density(df2$MPG.avg))
qqnorm(df2$MPG.avg, pch = 1, frame = FALSE)
qqline(df2$MPG.avg, col = "steelblue", lwd = 2)

## boxcox
# p1 <- powerTransform(df2$MPG.avg)
# df2$MPG.avg <- bcPower(df2$MPG.avg, p1$lambda)

# plot again
# plot(density(df$MPG.avg))
# qqnorm(df$MPG.avg, pch = 1, frame = FALSE)
# qqline(df$MPG.avg, col = "steelblue", lwd = 2)

```



check the Pairwise Pearson Correlations
From the Person Correlation matrix, there appears to be a significant amount of correlated relations between the predictor variables. It will thus be necessary to ensure that multicollinearity can be
a concern later in my model.
```{r message = FALSE, out.width = "40%"}
df %>%
    na.omit() %>%
    mutate(Cylinders = as.numeric(.$Cylinders)) %>%
    subset(select = -c(MPG.city, MPG.highway)) %>%
    # remove the non-numerical variables
    # https://bit.ly/3ZXRAMH
    .[, colnames(.)[!grepl("factor|logical|character", sapply(., class))]] %>%
    cor(., ) %>%
    round(., 2) %>%
    melt() %>%
    ggplot(., aes(x = Var1, y = Var2, fill = value)) +
        geom_tile() +
        scale_fill_distiller(direction = +1) +
        geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
        theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Linear Models to grasp the trend
```{r message = FALSE}
# simple linear model
# I use my intuition for now. Simply, I suppose the size of a car and engine and fuel tank are correlated with the car efficiency
simple <- lm(MPG.avg ~ Weight + Width + Length +
                    Fuel.tank.capacity + Horsepower,
                    data = df
                    )
# it is already high R-squared score. Width and Horsepower are insignificant. there can be multicollinearity among them.
summary(simple)
# Regression Diagnostics
simple.infl <- influence.measures(simple)
print(simple.infl)
influence.measures
# check the vif. In fact, Weight has the most highest vif
vif(simple)

# get rid of the weight parameter
simple_v2 <- lm(MPG.avg ~ Width + Length +
                    Fuel.tank.capacity + Horsepower,
                    data = df
                    )
# Indeed, the only one parameter became significant. This model is not preferrable in terms of Interpretability.
summary(simple_v2)
vif(simple_v2)
```

stepwise with AIC and BIC
```{r message = FALSE}
# use stepwise methods based on AIC and BIC
# df2 <- df %>%
#     subset(select = -c(MPG.city, MPG.highway, Min.Price, Max.Price, Manufacturer, Model)) %>%
#     na.omit() %>%
#     mutate(Cylinders = as.numeric(.$Cylinders)) %>%
#     .[, colnames(.)[!grepl("factor|logical|character", sapply(., class))]]
# m0 <- lm(data = df2, MPG.avg ~ Weight)
m1 <- lm(data = df2, MPG.avg ~ .)
M1_BIC <- stepwise(m1, direction = "forward/backward", criterion = "BIC")
M1_AIC <- stepwise(m1, direction = "forward/backward", criterion = "AIC")
summary(M1_BIC)
summary(M1_AIC)


# vif
vif(M1)
# since Weight's vif is higher than 10 so I will try to remove some of variables while maintaing R-squared score, and found that getting rid of Fuel.tank.capacity
m2 <- lm(data = df2, MPG.avg ~ Price + Cylinders + Wheelbase + Weight)
M2 <- stepwise(m2, direction = "backward", criterion = "BIC")


```

ANOVA for Origin, Passengers, Type
```{r message = FALSE, out.width = "50%"}


```

