---
title: "Regression for Causal Inference"
author: "Satoshi Ido"
output: pdf_document
header-includes:
  - "\\usepackage{listings}"
  - "\\lstset{breaklines=true,breakatwhitespace=true}"
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  tidy = TRUE
)
```

```{r}
library("ggplot2")
library("tidyverse")
library("MASS")
library("broom")
```

# Data
```{r warning=FALSE, message=FALSE}
email_data <- read.csv("/Users/satoshiido/Documents/programming/statistical-analysis/causal_inference/data/Kevin_Hillstrom_MineThatData_E-MailAnalytics_DataMiningChallenge_2008.03.20.csv")
head(email_data)
```

```{r warning=FALSE, message=FALSE}
# create the data w/o the womens E-Mail campaign
male_df <- email_data %>% 
  filter(segment != "Womens E-Mail") %>%
  mutate(treatment = if_else(segment == "Mens E-Mail", 1, 0))

# create the selection biased data
## set the seed
set.seed(1)

## make half depending on the condition
obs_rate_c <- 0.5
obs_rate_t <- 0.5

## create the biased data
biased_data <- male_df %>%
  mutate(
    obs_rate_c = if_else(
      (history > 300) | (recency < 6) | (channel == "Multichannel"), obs_rate_c, 1),
    obs_rate_t = if_else(
      (history > 300) | (recency < 6) | (channel == "Multichannel"), 1, obs_rate_t),
    random_number = runif(n = NROW(male_df))) %>%
    filter((treatment == 0 & random_number < obs_rate_c) | (treatment == 1 & random_number < obs_rate_t))

head(biased_data, 20)
```

# Regression
```{r warning=FALSE, message=FALSE}
# regression
biased_reg <- lm(data = biased_data, formula = spend ~ treatment + recency + history)
summary(biased_reg)
biased_reg <- lm(data = biased_data, formula = spend ~ treatment + history)
summary(biased_reg)
```

```{r warning=FALSE, message=FALSE}
# check only for treatment coefficient
biased_reg_coef <- tidy(biased_reg)
biased_reg_coef
```

# Biases in regression
```{r warning=FALSE, message=FALSE}
# simple regression with RCT data
rct_reg <- lm(data = male_df, formula = spend ~ treatment)
rct_reg_coef <- summary(rct_reg) %>% tidy()

# simple regression with biased data
nonrct_reg <- lm(data = biased_data, formula = spend ~ treatment)
nonrct_reg_coef <- summary(nonrct_reg) %>% tidy()

rct_reg_coef; nonrct_reg_coef
```

Regression with some covariates
```{r warning=FALSE, message=FALSE}
nonrct_mreg <- lm(data = biased_data, formula = spend ~ treatment + recency + channel + history)
nonrct_mreg_coef <- summary(nonrct_mreg)
# suppress selection bias a little by controlling for covariates
nonrct_mreg_coef
```

## OVB
```{r warning=FALSE, message=FALSE, include=FALSE}
# first model -> model A (alpha_i), second model -> # model B (beta_i), third model -> model C (gamma_i)
formula_vec <- c(spend ~ treatment + recency + channel, 
                 spend ~ treatment + recency + channel + history,
                 history ~ treatment + channel + recency)
names(formula_vec) <- paste("reg", LETTERS[1:3], sep = "_")

# enframe the vector
models <- formula_vec %>% enframe(name = "model_index", value = "formula")

# map the regression
df_models <- models %>%
  mutate(model = map(.x = formula, .f = lm, data = biased_data)) %>%
  mutate(lm_result = map(.x = model, .f = tidy))

# unnest the result
df_results <- df_models %>%
  mutate(formula = as.character(formula)) %>%
  dplyr::select(formula, model_index, lm_result) %>%
  unnest(cols = c(lm_result))
```

```{r warning=FALSE, message=FALSE}
# pull `treatment` parameters from the model A, B, C
treatment_coef <- df_results %>%
  filter(term == "treatment") %>%
  pull(estimate)

# pull `history` parameters from the model B
history_coef <- df_results %>%
  filter(model_index == "reg_B", term == "history") %>%
  pull(estimate)

# check OVB (beta_4 * gamma_1)
OVB <- history_coef * treatment_coef[3]
coef_gap <- treatment_coef[1] - treatment_coef[2]
OVB; coef_gap
```

## Post treatment bias
```{r warning=FALSE, message=FALSE}
# add the non-recommended variable to the model
cor_visti_treatment <- lm(data = biased_data, formula = treatment ~ visit + recency + history + channel) %>% tidy()
cor_visti_treatment
```

```{r warning=FALSE, message=FALSE}
bad_control_reg <- lm(data = biased_data, formula = spend ~ treatment + channel + recency + history + visit) %>% tidy()
bad_control_reg
```


# Regression EDA with vouchers data
```{r warning=FALSE, message=FALSE}
remotes::install_github("itamarcaspi/experimentdatar")
library("experimentdatar")
data(vouchers)
vouchers
```

```{r warning=FALSE, message=FALSE}
# prepare the regression
## prepare the character vectors for the regression
formula_x_base <- "VOUCH0"
formula_x_covariate <- "SVY + HSVISIT + AGE + STRATA1 + STRATA2 + STRATA3 + STRATA4 + STRATA5 + STRATA6 + STRATAMS + D1993 + D1995 + D1997 + DMONTH1 + DMONTH2 + DMONTH3 + DMONTH4 + DMONTH5 + DMONTH6 + DMONTH7 + DMONTH8 + DMONTH9 + DMONTH10 + DMONTH11 + DMONTH12 + SEX2"
formula_y <- c("TOTSCYRS","INSCHL","PRSCH_C","USNGSCH","PRSCHA_1","FINISH6","FINISH7","FINISH8","REPT6","REPT","NREPT",
               "MARRIED","HASCHILD","HOURSUM","WORKING3")

## create the simple regression formulas for each element in formula_y without covariates
base_reg_formula <- paste(formula_y, "~", formula_x_base)
names(base_reg_formula) <- paste(formula_y, "base", sep = "_")

## create the multiple regression formulas for each element in formula_y with covariates
covariate_reg_formula <- paste(formula_y, "~", formula_x_base, "+", formula_x_covariate)
names(covariate_reg_formula) <- paste(formula_y, "covariate", sep = "_")

## create the vectors for the models
table3_formula <- c(base_reg_formula, covariate_reg_formula)

## enframe the vectors
models <- table3_formula %>% enframe(name = "model_index", value = "formula")

# map the regression
## extract the data
regression_data <- vouchers %>% filter(TAB3SMPL == 1, BOG95SMP == 1)

df_models <- models %>%
  mutate(model = map(.x = formula, .f = lm, data = regression_data)) %>%
  mutate(lm_result = map(.x = model, .f = tidy))

# unnest the result
df_results <- df_models %>%
  mutate(formula = as.character(formula)) %>%
  dplyr::select(formula, model_index, lm_result) %>%
  unnest(cols = c(lm_result))
```

## Analysis of private school attendance and use of vouchers
Was the voucher used and did it help private school enrollment?
```{r warning=FALSE, message=FALSE}
# enrollment and voucher usage (PRSCHA_1 = if the student used the voucher, USNGSCH = if the student enrolled the private school)
using_voucher_results <- df_results %>%
  filter(term == "VOUCH0", str_detect(model_index, "PRSCHA_1|USNGSCH")) %>%
  dplyr::select(model_index, term, estimate, std.error, p.value)  %>%
  arrange(model_index)
using_voucher_results

# plot the result
using_voucher_results %>%
  ggplot(aes(y = estimate, x = model_index)) +
  geom_point() +
  geom_errorbar(aes(ymax = estimate + std.error*1.96,
                    ymin = estimate - std.error*1.96,
                    width = 0.1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,1, "cm"))
```

## Did voucher help the students to finish high school?
ggplot
```{r warning=FALSE, message=FALSE}
# extract the effect of VOUCH0 on PRSCH_C, INSCHL, FINISH6-8, REPT
## PRSCH_C = show if the student still attend the private school after 3 yrs, 
## INSCHL = show if the student still attend the school (public or pirvate) after 3 yrs, 
## FINISH6-8 = show if the student was retained in the 6th grade.
## REPT = show if the student was retained at least once before the survey
going_private_results <- df_results %>%
  filter(term == "VOUCH0",
         str_detect(model_index, "PRSCH_C|INSCHL|FINISH|REPT")) %>%
  dplyr::select(model_index, term, estimate, std.error, p.value) %>%
  arrange(model_index)

going_private_results %>%
  filter(str_detect(model_index, "covariate")) %>%
  ggplot(aes(y = estimate, x = model_index)) +
  geom_point() +
  geom_errorbar(aes(ymax = estimate + std.error*1.96,
                    ymin = estimate - std.error*1.96,
                    width = 0.1)) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,1, "cm"))
```

## Difference in Effectiveness by Gender
Replicate Angrist (2002) Table 4 & 6 bogota 1995
```{r warning=FALSE, message=FALSE}
# create the data for table 4
data_tbl4_bog95 <- vouchers %>%
  filter(BOG95SMP == 1, TAB3SMPL == 1,
         !is.na(SCYFNSH), !is.na(FINISH6), !is.na(PRSCHA_1), 
         !is.na(REPT6), !is.na(NREPT), !is.na(INSCHL), 
         !is.na(FINISH7),!is.na(PRSCH_C), !is.na(FINISH8), !is.na(PRSCHA_2), 
         !is.na(TOTSCYRS), !is.na(REPT)
         ) %>%
  dplyr::select(VOUCH0, SVY, HSVISIT, DJAMUNDI, PHONE, AGE,
         STRATA1:STRATA6, STRATAMS, DBOGOTA, D1993, D1995, D1997,
         DMONTH1:DMONTH12, SEX_MISS, FINISH6, FINISH7, FINISH8,
         REPT6, REPT, NREPT, SEX2, TOTSCYRS, MARRIED, HASCHILD,
         HOURSUM,WORKING3, INSCHL,PRSCH_C,USNGSCH,PRSCHA_1)
```

Women data
```{r warning=FALSE, message=FALSE}
# exctract women data
regression_data <- data_tbl4_bog95 %>% filter(SEX2 == 0)

# run the regression all together
df_models <- models %>%
  mutate(model = map(.x = formula, .f = lm, data = regression_data)) %>%
  mutate(lm_result = map(.x = model, .f = tidy))

# format the result
df_results_female <- df_models %>%
  mutate(formula = as.character(formula), gender = "female") %>%
  dplyr::select(formula, model_index, lm_result, gender) %>%
  unnest(cols = c(lm_result))
```

Men data
```{r warning=FALSE, message=FALSE}
# extract the men data
regression_data <- data_tbl4_bog95 %>% filter(SEX2 == 1)

# run the regression all together
df_models <- models %>%
  mutate(model = map(.x = formula, .f = lm, data = regression_data)) %>%
  mutate(lm_result = map(.x = model, .f = tidy))

# format the result
df_results_male <- df_models %>%
  mutate(formula = as.character(formula), gender = "male") %>%
  dplyr::select(formula, model_index, lm_result, gender) %>%
  unnest(cols = c(lm_result))
```

### Analysis of private school attendance and use of vouchers by gender
```{r warning=FALSE, message=FALSE}
# visualize analysis results to school attendance trends
## extract results for PRSCHA_1, USNGSCH
using_voucher_results_gender <- rbind(df_results_male, df_results_female) %>%
  filter(term == "VOUCH0", str_detect(model_index, "PRSCHA_1|USNGSCH")) %>%
  dplyr::select(gender, model_index, term, estimate, std.error, p.value) %>%
  # reorder the outputed dataframe
  arrange(gender, model_index) %>%
  filter(str_detect(model_index, "covariate"))

## ggplot
using_voucher_results_gender %>%
  filter(str_detect(model_index, "covariate")) %>%
  ggplot(aes(y = estimate, x = model_index)) +
  geom_point() +
  geom_errorbar(aes(ymax = estimate + std.error * 1.96,
                    ymin = estimate - std.error * 1.96,
                    width = 0.1)) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,1, "cm")) +
  facet_grid(gender ~ .)
```


### How Did voucher affect the students by gender to finish high school?
```{r warning=FALSE, message=FALSE}
# visualize analysis results to retention and years of school attendance
## extract the results for PRSCH_C,INSCHL,REPT,TOTSCYRS,FINISH
going_private_results_gender <- rbind(df_results_male, df_results_female) %>%
  filter(term == "VOUCH0",
         str_detect(model_index, "PRSCH_C|INSCHL|REPT|TOTSCYRS|FINISH")) %>%
  dplyr::select(gender, model_index, term, estimate, std.error, p.value) %>%
  arrange(model_index)

## ggplot
going_private_results_gender %>%
  filter(str_detect(model_index, "covariate")) %>%
  ggplot(aes(y = estimate, x = model_index)) +
  geom_point() +
  geom_errorbar(aes(ymax = estimate + std.error*1.96,
                    ymin = estimate - std.error*1.96,
                    width = 0.1)) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,1, "cm")) +
  facet_grid(gender ~ .)

# explore other factors because the above analysis showed that girls' academic persistence is less correlated with academic achievement, retention, and winning vouchers
# visualize the results of the analysis against HOUR
## extract the results of the analysis against HOUR
working_hour_results_gender <- rbind(df_results_male, df_results_female) %>%
  filter(term == "VOUCH0", str_detect(model_index, "HOUR")) %>%
  dplyr::select(gender, model_index, term, estimate, std.error, p.value) %>%
  arrange(gender, model_index)

## ggplot
working_hour_results_gender %>%
  filter(str_detect(model_index, "covariate")) %>%
  ggplot(aes(y = estimate, x = model_index)) +
  geom_point() +
  geom_errorbar(aes(ymax = estimate + std.error*1.96,
                    ymin = estimate - std.error*1.96,
                    width = 0.1)) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,1, "cm")) +
  facet_grid(. ~ gender)
```