---
title: "Bayesian Data Analysis – Assignment #1"
author: Youha Shin
date: "September 10, 2023"
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(reshape2)
library(moments)
```

```{r, include = FALSE}
setwd("/Users/youhashin/Library/CloudStorage/OneDrive-purdue.edu/Documents/STAT 656 [Bayesian Data Analysis]/HW/hw1")
# clear workspace
rm(list = ls())
```

<hr style="border: 1px solid gold;">

# Act I: Synthetic Data

## Question 1
**The Goal:** To determine the log-likelihood function $\log L(\rho, \sigma^2 | y_0, y_1, ..., y_n)$ for $(\rho, \sigma^2)^T$ 
for the $AR(1)$ model.  

*** 
  
1. We start by writing out the joint probability density function of the observations:  
$$p(y_0, y_1, ..., y_n | \rho, \sigma^2) = \prod_{i=1}^np(y_i | y_{i-1}, \rho, \sigma^2)$$   

    **Meaning:** This represents the probability of observing the entire sequence of observations $y_0, y_1, ..., y_n$ given the parameters $\rho$ and $\sigma^2$.
    This essentially sets the stage for Maximum Likelihood Estimation (MLE). 

  

2. Since the errors are assumed to be independent, we can write the conditional probability density function
of each observation as:  
$$p(y_i | y_{i-1}, \rho, \sigma^2) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{1}{2\sigma^2}(y_i-\rho y_{i-1})^2\right]$$  

    **Meaning:** The conditional probability density function represents the probability of observing $y_i$ given the previous observation $y_{i-1}$ and the parameters $\rho$ and $\sigma^2$.
    This step leverages the autoregressive nature of the model, allowing us to decompose the joint pdf into simpler conditional pdfs; this simplification is essentially for
    practical computations. 

3. Substituting this into the joint probability density function, we simplify it into a form that can be easily manipulated: 
$$p(y_0, y_1, ..., y_n | \rho, \sigma^2) = \left(\frac{1}{\sqrt{2\pi\sigma^2}}\right)^n \exp\left[-\frac{1}{2\sigma^2}\sum_{i=1}^n(y_i-\rho y_{i-1})^2\right]$$  
  
    **Meaning:** This step takes us closer to the function of the parameters $\rho$ and $\sigma^2$ that we can optimize. It integrates all pieces of the model
    into a single expression that captures the probability of the data given the parameters. 

4. Taking the logarithm of the joint probability density function, we attain the log-likelihood function in question:
$$\log L(\rho, \sigma^2 | y_0, y_1, ..., y_n) = -\frac{n}{2} \log(2\pi) -\frac{n}{2}\log(\sigma^2)-\frac{1}{2\sigma^2}\sum_{i=1}^n(y_i-\rho y_{i-1})^2$$

We may note that the first term in the log-likelihood function is a constant that does not depend on the parameters, so it can be ignored
when maximizing the likelihood. The second term is proportional to the log of the variance, so maximizing the likelihood with respect to $\sigma^2$ is equivalent
to minimizing the second term. The third term is proportional to the sum of squared errors, so maximizing the likelihood with respect to $\rho$ is equivalent is 
equivalent to minimizing the third term.   
  
In sum, maximizing this function gives the estimates of $\rho$ and $\sigma^2$ that best fit the data. 

<hr style="border: 1px solid black;">

## Question 2
**The Goal:** Write an `R` function that computes the log of the likelihood functions
for $(\rho, \log(\sigma))^T$ for the dataset provided. Furthermore, the task includes creating a contour plot visualization of the log-likelihood function.

*** 



```{r} 
## import data
set.seed(1) # set seed for reproducibility

q2 <- read.csv("computation_data_hw_1.csv", header = TRUE)

y <- q2$x

## Define the log-likelihood function
ar_loglik <- function(rho, log_sig) {

    # Get sigma from log calculation
    sig <- exp(log_sig)

    ## Initialize log-likelihood
    loglik <- 0

    ## Compute log-likelihood based on AR(1) model
    n <- length(y)
    loglik <- -n/2 * log(2 * pi * sig^2)
    sum_term <- sum((y[2:n] - rho * y[1:(n-1)])^2)
    loglik <- loglik - sum_term / (2 * sig^2)

    return(loglik)
}

# Creating grid of rho and log_sigma values
rho_values <- seq(-0.99, 0.99, by = 0.05) # rho must be between -1 and 1
log_sig_values <- seq(-2, 2, by = 0.05) # log(sigma) can be any value
# we will use a smaller range for illustration purposes

# Compute the log-likelihood for each combination using the 'outer' function
z <- outer(rho_values, log_sig_values, 
            Vectorize(function(rho, log_sig) ar_loglik(rho, log_sig)))

# Create the contour plot
q1_plot <- contour(rho_values, log_sig_values, z, 
        xlab = expression(rho),
        ylab = expression(log(sigma)),
        main = "Log-Likelihood Contour Plot for AR(1) Model",
        levels = pretty(range(z), 20),
        col = rainbow(10)
        )

```

```{r} 
## Attempt to zoom in
rho <- seq(0, 1, .01)
log_sig <- log(seq(.4, 1, .01))

loglike <- outer(rho, log_sig, 
            Vectorize(function(rho, log_sig) ar_loglik(rho, log_sig)))

contour(rho, log_sig, loglike, 
        xlab = expression(rho),
        ylab = expression(log(sigma)),
        main = "Log-Likelihood Contour Plot for AR(1) Model",
        levels = pretty(range(loglike), 20),
        col = rainbow(10)
        )

```

``` {r}
### Attempt to zoom in further
rho <- seq(0.4, .6, .001)
log_sig <- log(seq(.45, .55, .001))

loglike <- outer(rho, log_sig, 
            Vectorize(function(rho, log_sig) ar_loglik(rho, log_sig)))

contour(rho, log_sig, loglike, nlevels = 50, 
        xlab = expression(rho),
        ylab = expression(log(sigma)),
        main = "Log-Likelihood Contour Plot for AR(1) Model",
        levels = pretty(range(loglike), 20),
        col = rainbow(10)
        )

```


### Interpretation of the Contour Plot
The log-likelihood function measures how well a particular set of parameters fits a given dataset. In this case, the 
parameters are $\rho$ and $\sigma^2$, which are the autoregressive coefficient and the standard deviation of the `AR(1)` model, respecitvely. 
The log-likelihood function is maximized when the parameters are set to the values that best fit the data.  
  
The resulting contour plot shows the values of the log-likelihood function for different combinations of $\rho$ and $\sigma$.
The highest point on the contour plot corresponds to the values of $\rho$ and $\sigma$ that maximize the log-likelihood function. 



<hr style="border: 1px solid black;">

## Question 3
**The Goal:** To write an `R` function that computes the log of the posterior density (up to a constant) for
$(\rho, \log(\sigma))^T$ under the prior specified by $\rho$ ~ $Uniform(-1, 1)$ and $\log(\sigma)$ ~ $N(0, 10^2)$ independently. 
Furthermore, the task includes creating a visualization of this function and providing interpretations of the results.

***

To compute the log of the posterior density for $(\rho, \log(\sigma))^T$, we utilize Bayes' Theorem:
$$ Posterior \propto Likelihood \times Prior$$
  
The likelihood is already known from the `AR(1)` model and given the specified prior distributions for 
$\rho$ and $\log(\sigma)$, we can compute the log posterior density (up to a constant). 

```{r}
# Function to compute log-posterior density
compute_log_posterior <- function(rho, log_sig, y) {

    # Likelihood computation based on AR(1) model
    n <- length(y)
    sig <- exp(log_sig)
    log_likelihood <- -n/2 * log(2 * pi * sig^2) - sum((y[2:n] - rho * y[1:(n-1)])^2 / (2 * sig^2))

    # Prior computation
    # For rho ~ Uniform(-1, 1)
    if(rho >= -1 && rho <= 1) { 
        log_prior_rho <- -log(2) # log(1/2) = -log(2)
    } else {
        log_prior_rho <- -Inf # log(0) = -Inf
    }

    # For log(sigma) ~ N(0, 10^2)
    log_prior_log_sig <- -0.5 * (log_sig^2) / (100) # ignoring constant terms in Gaussian pdf
    # dnorm(log_sig, 0, 10, log=T) # alternative way to compute log_prior_log_sig

    # Compute log-posterior density (up to a constant)
    log_posterior <- log_likelihood + log_prior_rho + log_prior_log_sig

    return(log_posterior)
}

# Compute log-posterior values on the grid
rho_values <- seq(-1, 1, length.out = 100) # rho must be between -1 and 1
log_sig_values <- seq(-2, 2, length.out = 100) # log(sigma) can be any value
log_posterior_values <- matrix(0, nrow = 100, ncol = 100) # initialize matrix of log-posterior values

for (i in 1:100) {
    for(j in 1:100) {
        
        # Compute log-posterior values for each combination of rho and log(sigma)
        log_posterior_values[i, j] <- compute_log_posterior(rho_values[i], log_sig_values[j], y)
    }
}

# Melt the matrix to a data frame for ggplot
df <- melt(log_posterior_values)
colnames(df) <- c("Rho", "LogSigma", "LogPosterior")

# Create the contour plot
q3_plot <- ggplot(df, aes(x = rho_values[Rho], y = log_sig_values[LogSigma], z = LogPosterior)) +
        geom_tile(aes(fill = LogPosterior)) + # fill the tiles with log-posterior values
        geom_contour(aes(z = LogPosterior), color = 'white') + # add contour lines
        scale_fill_gradientn(colors = viridis::viridis(7)) + # color gradient
        labs(x = expression(rho), y = expression(log(sigma)), fill = 'Log-Posterior Value') +
        theme_minimal() + # minimal theme
        ggtitle('Log-Posterior Density Contour Plot')

q3_plot
```

#### How does this function correspond to the log-likelihood function?
The log-posterior combines the log-likelihood (which measures how well the model explains the data) and the 
log-priors (which encapsulate our prior beliefs about the parameters). In the resulting plot, one can observe that
areas of higher density reflect parameter combinations that are more consistent with both the data and our prior beliefs.
When examining the visualization, the highest values of the log-posterior density appear to be concentrated around $\rho = 0.5$ and $\log(\sigma) = 0$.  
    

#### Is the prior specification overly informative?
1. **For $\rho:$** The $Uniform(-1, 1)$ prior is rather uninformative within its bounds, giving equal 
weight to all values in this range. However, it does strictly confine $\rho$ to this range, excluding values 
outside of it entirely.  
2. **For $\log(\sigma):$** The Normal prior with a variance of $10^2$ tells us that values of $\log(\sigma)$ close to 0
are more likely than values that are far from 0. However, the prior is quite broad, so it does not strongly favor any particular value of $\log(\sigma)$.  

Given the wide range for $\log(\sigma)$ and the flatness of the Uniform prior for $\rho$ within its bounds,
the prior specification here does not seem overly informative. This allows the data to play a significant role in 
determining the posterior distribution, as reflected in the contour plot.  

<hr style="border: 1px solid black;">

## Question 5
**The Goal:** Draw 1,000 values of $\rho, \log(\sigma))^T$ from a discrete grid approximation to the posterior. Describe your
choice of discrete grid.  

***
To attain the Goal, we may leverage the work done in calculating the log-posterior values over the grid. 
Specifically, we can use these values to create a probability distribution over the grid points, which can 
then be sampled to approximate the posterior distribution.  
  
The protocol is as follows:  
  
1. **Normalize the Log-Posterior Values:** Convert log-posterior values to posterior probabilities by exponentiating them 
and then normalizing so they sum to one; this allows us to interpret the values as probabilities.
$$P(\rho, \log(\sigma) | data) \propto \exp(log posterior)$$  

2. **Create a Joint Grid:** Pair each $\rho$ value with each $\log(\sigma)$ value along its corresponding
normalized posterior probability.  

3. **Sampling:** Use the `sample` function to draw 1,000 samples from this discrete distribution. 


```{r}
## Step 1: Normalize the Log-Posterior Values
# We first exponentiate the log-posterior values and then normalize them so they sum to one
exp_log_posterior_values <- exp(log_posterior_values - max(log_posterior_values)) 
# Note: subtracting the max value ensures that the largest value is 0, which prevents overflow

normalized_posterior_values <- exp_log_posterior_values / sum(exp_log_posterior_values) # normalize

## Step 2: Create a Joint Grid
grid_data <- expand.grid(rho = rho_values, log_sigma = log_sig_values) # create a joint grid
grid_data$posterior_prob <- as.vector(normalized_posterior_values) # add posterior probabilities to the grid

## Step 3: Sample 1000 values from this discrete distribution
set.seed(42) # set seed for reproducibility
sample_indices <- sample(1:nrow(grid_data), size = 1000, replace = TRUE, prob = grid_data$posterior_prob) # sample indices
sampled_values <- grid_data[sample_indices, ] # get sampled values

# The sampled_values df now contains 1000 samples of (rho, log(sigma)) from the posterior distribution
head(sampled_values)
```

#### Choice of Discrete Grid
The choice of the grid size matters as it represents the a key tradeoff between accuracy of approximation and computational efficiency.
The grid should be 'fine' enough to capture the main features of the posterior distribution, 
but not so fine that it becomes computationally expensive to sample from.  
  
In this case, the discrete grid is created using 100 values for $\rho$ and 100 values for $\log(\sigma)$, resulting in a total of 10,000 grid points,
which should be sufficient to capture the main features of the posterior distribution.  

Ultimately, the grid was chosen based on the contour plot of the log-posterior to ensure it spans the regions where the posterior density is significant. 
The resulting `sampled_values` data frame contains 1,000 samples of $(\rho, \log(\sigma))^T$ from the posterior distribution. These samples can be used to 
estimate the posterior, mean, variance, and other properties of the distribution.  

<hr style="border: 1px solid black;">

## Question 7
**The Goal:** To use the sample draws to calculate the following summaries for each parameter: 0.025, 0.25, 0.5, 0.75, 
0.975 quantiles, mean, standard deviation, skewness, and kurtosis. 

***

```{r}
# Create an empty list to store summaries
summaries <- list()

# Loop through each variable
for (var in c("rho", "log_sigma")) {

    # Extract the samples for the variable
    samples <- sampled_values[[var]]

    # Calculate the summaries
    summaries[[var]] <- list(
        quantile_0.025 = quantile(samples, 0.025),
        quantile_0.25 = quantile(samples, 0.25), 
        quantile_0.5 = quantile(samples, 0.5),
        quantile_0.75 = quantile(samples, 0.75), 
        quantile_0.975 = quantile(samples, 0.975),
        mean = mean(samples), 
        std_dev = sd(samples), 
        skewness = skewness(samples),
        kurtosis = kurtosis(samples)
    )
}

# View the summaries
print(summaries)
```

<hr style="border: 1px black;">

## Question 8
**The Goal:** Write an `R` function that takes parameters $(\rho, \log(\sigma))^T$ and simulates a new
dataset $y^{rep}$ according to the `AR` process. Use the two earlier `R` functions to generate 1,000 posterior
predictive samples; summarize the draws. 

***

```{r}
## Step 1: Define a function to simulate y_rep

simulate_ar_proc <- function(rho, log_sigma, n) {
    sigma <- exp(log_sigma) # get sigma from input
    y_rep <- numeric(n) # initialize vector to store simulated values
    y_rep[1] <- rnorm(1, mean = 0, sd = sigma) # simulate first value from N(0, sigma^2)

    # Simulate the remaining values
    for (i in 2:n) {
        y_rep[i] <- rho * y_rep[i-1] + rnorm(1, mean = 0, sd = sigma) # simulate from N(rho * y[i-1], sigma^2)
    }

    return(y_rep)
}

## Step 2: Generate 1000 posterior predictive samples

n <- length(y) # length of OG dataset
posterior_predictive_samples <- matrix(0, nrow = 1000, ncol = n) # initialize matrix to store posterior predictive samples

for (i in 1:1000) {
    sampled_row <- sampled_values[i, ] # get the sampled values for each row
    rho_sample <- sampled_row$rho # get rho sample
    log_sigma_sample <- sampled_row$log_sigma # get log(sigma) sample

    # Simulate a new dataset using the sampled values
    posterior_predictive_samples[i, ] <- simulate_ar_proc(rho_sample, log_sigma_sample, n)    
}

## Step 3: Summarize the draws
summaries <- data.frame(matrix(0, nrow = 1000, ncol = 9)) # initialize matrix to store summaries
colnames(summaries) <- c("quantile_0.025", "quantile_0.25", "quantile_0.5", "quantile_0.75", "quantile_0.975", 
                            "mean", "std_dev", "skewness", "kurtosis")


summaries$mean <- apply(posterior_predictive_samples, 1, mean)
summaries$std_dev <- apply(posterior_predictive_samples, 1, sd)
summaries$skewness <- apply(posterior_predictive_samples, 1, skewness)
summaries$kurtosis <- apply(posterior_predictive_samples, 1, kurtosis)
summaries[ , 1:5] <- apply(posterior_predictive_samples, 1, quantile, probs = c(0.025, 0.25, 0.5, 0.75, 0.975))

# View the summaries
head(summaries)

```


<hr style="border: 1px solid black;">
## Question 9
**The Goal:** Compare the observed data to the posterior predictive summaries; create a plot where the 
observed data is superimposed on the posterior predictive trajectories. Interpret the model fit and its aptness.

***

```{r}
## Combine posterior predictive samples and observed data
combined_data <- data.frame(
    Time = rep(1:n, 1000), # time
    Value = c(as.vector(posterior_predictive_samples), rep(y, 1000)), # posterior predictive samples + observed data
    Type = rep(c(rep("Predicted", nrow(posterior_predictive_samples)), rep("Observed", nrow(posterior_predictive_samples))), each = n) # type of data
)

## Plot the data
q9_plot <- ggplot(combined_data, aes(x = Time, y = Value, color = Type)) +
        geom_line() +
        labs(x = "Time", y = "Value", color = "Type") +
        theme_minimal() +
        scale_color_manual(values = c("Predicted" = "blue", "Observed" = "red")) +
        ggtitle("Posterior Predictive Samples vs. Observed Data")

q9_plot


```


<hr style="border: 1px solid gold;">

# Act II: Real Data 

## Question 1
**The Goal:** To assess whether the information given on Github is sufficient for analyzing the data; provide 
justifications for your case. If determined not to be sufficient, address what other information you would have requested/collected. 

***

The primary factors we will use to assess the sufficiency of information are: Data Quality, Data Scope, and Documentation.  
  
#### On Data Quality: 
The accuracy, completeness, consistency, and granularity are all integral for proper analysis. 
The Github page mentions that the data are from reputable sources, including the CDC and federal government, which 
suggests that the data are likely to be of high quality. Moreover, the NYT acknowledges how their methodology is inherently
imperfect due to the lack of uniform reporting across states and counties, as well as errors in reporting; however, the organization
clearly has a strong team and commitment to data quality, thus we can assume that this represents one of the more robust data sources available for COVID-19.  
  
The datasets appear to have all the necessary columns present across its temporal range, and appears to be consistent in its structure. 

#### On Data Scope:
The `COVID-19 Data for the U.S.` overall contains columns for the date, number of cases (including both confirmed and probable), and number of deaths
(including both confirmed and probable). The temporal range begins on January 21, 2020 and ends on August 25, 2020. The `COVID-19 Data for Individual States` operates on the same timeline and
contains the same columns, but it also includes the state name and the fips code for each state.   
  
While all the necessary columns seem to be present for a basic time-series analysis, there is 
assuredly more information that could be useful for a more advanced analysis. If resources were unlimited, additional data regarding the following would significantly enhance the 
level of analysis possible:  
  
  1. Demographic Information: Age, gender, and other demographic data would provide deeper insights into which groups are most affected.
  2. Hospitalization Rates: This would be useful for understanding the severity of cases.
  3. Testing Rates: Knowing how many tests have been conducted would provide context to the number of cases and would enable us to make more targeted assumptions about the data.
  4. Vaccination Rates: A critical factor that could influence the number of new cases and deaths. 
  5. Seasonality: This would be informative for understanding the impact of the pandemic on different seasons and how the pandemic may evolve across time.
  6. Policy Data: Information on lockdowns, mask mandates, etc., could help in understanding public behavior and the potential spread of the virus.  
  
#### On Documentation:
The Github page provides a detailed description of their tracking process, helpful differentiation between the different types of data (historical vs. live), and a general overview of the methodology
and definitions, including the different types of cases and deaths.  
  
However, the documentation, while useful as a higher-order overview, does not include access to more detailed information, such as data dictionaries or codebooks to understand the 
variables and relationships with other variables. While this is perfectly understandable given the duress of the circumstances, it does introduce limitations in understanding the data on a more granular 
level and impacts some of the assumptions we may make about the data. Comprehensive metadata which explains the units, data types, and data sources would
drastically improve the reliability of any analysis; and further information on how the data was collected, including sampling techniques or adjustments made, would be integral for 
gaining clarity about the limitations at hand.  

<hr style="border: 1px solid black;">

## Question 2
An AR(1) model can be expressed as: 
$$ Y_t = \rho Y_{t-1} + \epsilon_t$$
where $\rho$ is the autoregressive coefficient and $\epsilon_t$ is the error term at time $t$.  
  
While using raw data can be straightforward and be reasonable if the AR assumptions are satisfied, it 
may not capture the underlying trends or seasonality in the data.  
  
In the case of the COVID-19 data, the raw data is the number of total cases and deaths (including both confirmed and probable).
This data is highly skewed, with a few days with zero cases and the later dates reporting very high numbers of cases and deaths. 
As such, this skewness makes it difficult to fit an `AR(1)` directly to the raw data, and we should consider appropriate transformations
based on the assumptions of the autoregressive model.  
  
- Stationarity: The series should be stationary, that is, its statistical properties (such as mean and variance) should remain constant
    over time. In the AR(1) model, a requirement for stationarity is that the absolute value of $\rho_1$ is less than 1. The COVID-19 data is very likely
    to be non-stationary due to trends and seasonality. Therefore, differencing the data at least once is generally a good protocol. 
- Independence: The errors are assumed to be independent and identically distributed; if there is any correlation between errors, the estimated of the AR model parameters
    will be inefficient. The COVID-19 data exhibits serial correlation, as the number of cases/deaths on any given day is influenced by the number of cases on the previous day.
    This strengthens the case to implement a first-order differencing of the time series. 
- Normality: The errors are assumed to be normally distributed; if the errors are not normally distributed, it may affect the validity of the model's predictions. This can be checked using
    Q-Q plot visualizations and statistical tests. 

In general, it is good practice to implement a log-transformation on the data to reduce the skewness and stabilize the variance. While this can make the interpretation of results more complex, given the 
scientific importance of accurate modeling over interpretability in this context, transformations are justifiable and prudent. In sum, a reasonable protocol would encompass checking for stationarity in the data,
and if necessary, differencing the data at least once. Thereafter, a log-transformation can be applied to the differenced data to reduce the skewness and stabilize the variance. These procedures would ultimately
make the errors more independent and will likely result in a better model fit. This approach aligns with the statistical assumptions of the autoregressive model and the nature of the pandemic data. 

```{r}
# Load the data
covid_us <- read.csv("covid_us.txt", header = TRUE, sep = ",")

# ar_mod1 <- arima(log(covid_us$cases), order = c(1, 1, 0))
ar_mod1 <- arima(covid_us$cases, order = c(1, 1, 0))
# check the residuals for normality
qqnorm(ar_mod1$residuals)
qqline(ar_mod1$residuals)

# check the residuals for autocorrelation
acf(ar_mod1$residuals)

# check the residuals for heteroskedasticity
plot(ar_mod1$residuals, type = "l")
abline(h=0, col = "red")

### Hyun: Taking the differences of the data
### covid_us$new_cases = c(0, covid_us[2:218,2] - covid_us[1:217,2])
### covid_us$newdeath = c(0, covid_us[2:218,3] - covid_us[1:217,3])

### ar_mod1 <- arima(covid_us$new_cases, order = c(1, 0, 0))
### ar_mod1 <- arima(log(covid_us$new_cases + 100), order = c(1, 0, 0)) # adding 100 to avoid log(0)
#### need to justfy +100; see dist of transformed value 


### new_death_rate
### covid_us$newdeathrate = (covid_us[,5]+3)/(covid_us[,4]+100)


# Take the logarithm of the data
# covid_us$log_cases <- log(covid_us$cases)

# Create a new column for the first difference
# covid_us$diff_log_cases <- (covid_us$log_cases - lag(covid_us$log_cases))

# Fit an AR(1) model to the first difference
# ar_mod2 <- arima(covid_us$diff_log_cases, order = c(1, 0, 0))

# check the residuals for normality
# qqnorm(ar_mod2$residuals)
# qqline(ar_mod2$residuals)

# check the residuals for autocorrelation
# acf(ar_mod2$residuals)

# check the residuals for heteroskedasticity
# plot(ar_mod2$residuals, type = "l")
# abline(h=0, col = "red")


```


<hr style="border: 1px solid black;">

## Question 3
**The Goal:** To specify a prior for all of the model parameters; justify your choice. 

***

#### Prior for the Autoregressive Coefficient $\rho$:
The autoregressive coefficient in an `AR(1)` model represents the correlation between consecutive observations. 
In the context of COVID-19 cases/deaths, it is logical that this coefficient is positive (since the number of cases/deaths
on one day is likely to be similar to the number of cases on the previous day. However, it is unlikely to be exactly 1 (implying
that the number of cases/deaths is exactly the same on consecutive days). Hence, a prior that assigns high probabilities to values
near 1 seems to be reasonable.  
  
In this context, we may use a Beta distribution to model the prior distribution of $\rho$. The Beta distribution is a continuous distribution
probability distribution that is defined on the interval (0, 1). The two parameters $\alpha$ and $\beta$ control the shape of the distribution, 
and it is noted that the Beta distribution is Uniform when $\alpha = \beta = 1$. If we believe that the coefficient is likely to be close to 1, 
we can choose larger $\alpha$ and $\beta$ values; alternatively, if we believe that the coefficient is likely to be close to 0, we can choose smaller
parameter values.  
  
Furthermore, the Beta distribution is a conjugate prior for the AR(1) model, which means that the posterior distribution of the autoregressive
coefficient can be easily computed using a Beta prior. This is a desirable property as it simplifies the mathematical convenience and computations.

#### Prior for the Standard Deviation of the Error Term $\sigma$:
The standard deviation of the noise term in an AR(1) model represents the variability in the number of cases that is not explained by the autoregressive process.
Given the multitude of factors that can influence the spread of COVID-19, it is reasonable to assume that the standard deviation is relatively large. As such, 
a prior that assigns a high probability to large values of $\sigma$ seems to be sensible approach.  
  
We may use the Inverse Gamma Distribution to model the prior distribution of $\sigma$. The Inverse Gamma Distribution is a continuous probability distribution defined on 
the interval (0, $\infty$). Once more, the two parameters $\alpha$ and $\beta$ control the shape of the distribution, and it is noted that the Inverse Gamma distribution is the
same as the Exponential distribution when $\alpha = 1$. If we believe that the standard deviation is likely to be large, we can choose $\alpha$ to be small and $\beta$ to be large.
On the other hand, if we believe that the standard deviation is more likely to be small, we can choose $\alpha$ to be large and $\beta$ to be small.  
  
As with the case of the Beta distribution, the Inverse Gamma is a conjugate prior for the AR(1) model. This also makes the inverse gamma prior a convenient and pragmatic choice
in this problem set. 