---
title: "STAT 527 HW 1"
author: "Satoshi Ido (ID: 34788706)"
date: 23 January 2023
output: pdf_document
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    warning = FALSE,
    message = FALSE)
```

1\
(10 points) According to Newton's universal law of gravitation, \
the acceleration of an object in the direction of the sun due to the sun's gravity can be written in the form \
$$
a = \frac{1}{r^2}
$$ \
Here, r is the distance of the object from the sun's center, in astronomical units (AU) of distance. 
One AU is the average distance of the Earth from the sun, about 150 million kilometers. 
The units of a are scaled for convenience in this version of Newton's equation so that one unit of acceleration is experienced at a distance of 1AU. 
Use the equation to calculate the gravitational accelerations at each of the planets' average distances from the sun:
```{r message = FALSE}
Mercury <- 1 / (0.39)^2
Venus <- 1 / (0.723)^2
Mars <- 1 / (1.524)^2
Jupiter <- 1 / (5.203)^2
Saturn <- 1 / (9.539)^2
Uranus <- 1 / (19.18)^2
Neptune <- 1 / (30.06)^2
Pluto <- 1 / (39.53)^2
a <- c(Mercury, Venus, Mars, Jupiter, Saturn, Uranus, Neptune, Pluto)
# the gravitational accelerations at each of the planets’ average distances from the sun
print(a)
```

2\
(10 points) Using the equation from Question 1, draw a graph of the gravitational acceleration (vertical axis) versus a range of values of r ranging from around 0.4AU (distance of Mercury) to around 5.2AU (distance of Jupiter). 
According to Newton's gravitational law, is there any distance at which the sun's gravity is escaped entirely?
```{r message = FALSE, out.width = "70%"}
r <- 0.4:5.2
a <- 1 / r^2
plot(r, a, xlab="AU", ylab="the gravitational acceleration", xaxp = c(0.4, 5.2, 8))
```

3\
(10 points) The following (next page) are the population sizes of the United States through its entire history, according to the U.S. Census. 
Construct a line plot (type=”I”) of the U.S. population (vertical axis) versus time (horizontal axis). 
By the way, rounding the population sizes to the nearest 100,000 or so will hardly affect the appearance of the graph.
Change the script and repeat the plot six more times (saving the graph each time), using type=”p”, type=”b”, type=”C”, type=”o”, type=”h”, and type=”l”. 
Compare the different graph types. What different aspects of the data are emphasized by the different graph types?
```{r message = FALSE, figure-side, fig.show="hold", out.width="48%"}
library("plyr")

data <- c(3929214, 5308483, 7239881, 9638453, 12860702, 17063353, 23191876, 31443321, 38558371, 50189209,
            62979766, 76212168, 92228496, 106021537, 123202624, 142164569, 161325798, 189323175, 213302031,
            236542199, 258709873, 291421906
        )
time <- seq(1790, 2000, 10)
round_data <- round_any(data, 100000)
# The line plot of of the U.S. population vs time
plot(time, round_data, type = "l", las = 1, cex.axis = 0.5)

# Compare with the different graph types
plot(time, round_data, type = "p", las = 1, cex.axis = 0.5)
plot(time, round_data, type = "b", las = 1, cex.axis = 0.5)
plot(time, round_data, type = "c", las = 1, cex.axis = 0.5)
plot(time, round_data, type = "o", las = 1, cex.axis = 0.5)
plot(time, round_data, type = "h", las = 1, cex.axis = 0.5)
```