---
title: "STAT 527 HW 5"
author: "Satoshi Ido (ID: 34788706)"
date: 20 February 2023
output: pdf_document
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    warning = FALSE,
    message = FALSE)
```

1. (10 points) Fibonacci sequence. If you put some coefficients in the recursion, like ri+1 = ari + bri-1,
where you choose the numbers a and b, then you will have an entirely new sequence. Write a function that takes a, b and the sequence length n as arguments. The function should return the entire sequence. Choose appropriate input arguments and plot the returned sequence.
```{r}

r <- function(a, b, n) {
    # start with i = 1 and i-1 = 0
    i <- 0; r1 <- 0; r2 <- 1
    while (i < n) {
        print(r1)
        r3 <- a * r2 + b * r1
        # update values
        i <- i + 1
        r1 <- r2
        r2 <- r3
    }
}
# Choose a = 1, b = 1, n = 10
r(1, 1, 10)
```

2. (20 points) 0-1 knapsack problem. Given a set of items, each with a weight and a value, determine which items to include in the collection so that the total weight is less than or equal to a given limit and the total value is as large as possible.
Mathematically, given a set of n items numbered from 1 up to n, each with a weight wi and a value vi, along with a maximum weight capacity W (assumed to be an integer), the 0-1 knapsack problem aims to solve the following optimization problem:
```{r}
wt <- c(10, 20, 30)
vt <- c(60, 100, 120)
w <- 50
n <- length(wt)
M <- matrix(0, nrow = n + 1, ncol = w)
k <- function(x) {x + 1}

# function
Knapsack <- function(weight, value, capacity) {
    # to skip the first row, start with 2 to k(n)
    for (i in 2:k(n)) {
        for (j in 1:w) {
    # if the weight of new item is over j-th, use the value at {(i-1), j}
            if (j < wt[i - 1]) {
                M[i, j] <- M[i - 1, j]
            }
    # ifelse, comparing m[i-1,w] & m[i-1,w-wi]+v and pick the bigger one
        else {
            M[i, j] <- max(M[i - 1, j], M[i - 1, j - wt[i - 1]] + vt[i - 1])
            }
        }
    }
    return(M)
}

# wt = {10,20,30}, vt = {60,100,120} and W = 50
M <- Knapsack(wt, vt, w)
# change the row names
rownames(M) <- 0:3

M
```