---
title: "STAT 527 HW 7"
author: "Satoshi Ido (ID: 34788706)"
date: 27 March 2023
output: pdf_document
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    warning = FALSE,
    message = FALSE)
```



```{r}
# data prep
library("lattice")
library("dplyr")
data(mtcars)
head(mtcars)
```

Q1. Plot the density plot of the Weight (1000 lbs) of the cars with the function densityplot().
```{r, out.width = "70%"}
# the density plot of Weight
attach(mtcars)
densityplot( ~ wt,
        xlab = "Weight (1000 lbs)",
        ylab = "Kernel Density",
        plot.points = FALSE
    )
```

Q2. Add a new column transmission to the data frame mtcars. The new column should be a factor vector, each element labeled Automatic or Manual according to the Transmission type (already given in one of the data columns).
```{r, out.width = "70%"}
mtcars <- transform(
                mtcars,
                transmission = ifelse(
                    am == 0, "Automatic", "Manual"
                )
            )
# change the data type from the character into the factor
mtcars$transmission <- as.factor(mtcars$transmission)
head(mtcars$transmission)
class(mtcars$transmission)
```

Q3. Use barchart() to compare the number of vehicles with the two different transmission types.
```{r, out.width = "70%"}
barchart(
        mtcars$transmission,
        horiz = TRUE,
        main = "the number of vehicles with different transmission types",
        xlab = "the number of vehicles",
        ylab = "transmission types",
        col = "grey"
    )
```

Q4. Use xyplot() to explore the relation between Miles per Gallon and Weight (1000 lbs), grouped by transmission.
```{r, out.width = "70%"}
xyplot(
    mpg ~ wt,
    data = mtcars,
    groups = transmission,
    xlab = "Weight (1000 lbs)",
    ylab = "Miles per Gallon",
    main = "the relation between Miles per Gallon and Weight (1000 lbs)"
    )
```

Q5. Use bwplot() to explore the distribution of Miles per Gallon, grouped by transmission
```{r, out.width = "70%"}
attach(mtcars)
bwplot(mpg ~ transmission,
    xlab = "Miles per Gallon",
    main = "the distribution of Miles per Gallon grouped by transmission"
    )
```